---
title: "Analytics 201: Session 3"
output:
  html_document
---

# Data Preprocessing (cont.)

```{r}
<<<<<<< HEAD

```


# Basic Plotting

```{r}

```

=======
```

# R studio "help"
If we know the name of a function, but are unsure on how to use it, we can always run the command __?function_name__ in R studios to pull up the help log documentation (e.g. ?select, ?hist). Note that if the function is a part of a package, the package must be imported first for the help log to work.

# Load Packages 
```{r}
library(dplyr)
library(ggplot2)
```

# Plotting
One key tool in analytics is data visualization. Indeed, a picture speaks a thousand words. We will first investigate some basic plots that base R (i.e. no packages) offers, then transition into using a powerful package called ggplot2 

## Data
We will be using a dataset called _mpg_, which is self-loaded by importing ggplot (make sure you call library(ggplot2) first). You can find more about the dataset [here](https://ggplot2.tidyverse.org/reference/mpg.html). 

```{r}
print(mpg)
```

## Some basic plots

### hist
If we want to see the distribution of city miles per gallon for all models(cty), we can make a histogram. To call up a column from a dataframe, we use the __$__ symbol. Note that it's sufficient to only specify the column graphed, but we can add in (optional) additional arguments to improve our graph.
```{r}
# hist(mpg$cty) this is sufficient
hist(mpg$cty, main = 'Histogram of cty', xlab = 'miles per gallon')
```
In general, some main "embellishments" that work in base R:

* main: the title of the graph
* ylab/xlab: the y and x labels
* ylim/xlim: the limits of the y/x axis. Note that they take in a vector with two numbers that signify the lower/upper limits. e.g. xlim = c(0, 100)

### pie
We may want to understand the distribution of car types using a pie chart. We first need to count how many instances of each type of car is present. There's a funciton in dplyr called count that does this. 
```{r}
car_types = mpg %>% count(class)
print(car_types)

```
Note that under the hood, this is the real function being run:
__mpg %>% group_by(class) %>% summarize(counts = n())__
i.e. first group by class, and then do a count for each class, with n() being the function for count

Then we can call up a pie chart. The format here is pie(counts of the car types, (optional)labels of car types, (optional)radius of the chart)
```{r}
pie(car_types$n, labels = car_types$class, radius = 1)
```

While base R does offer a variety of different types of plots (i.e. line, hist, scatter, etc), the amount of customization is limited. Furthermore, the syntax/pre-treatment required can get complicated especially when we try to have multiple components in the same graph (e.g. line charts with multiple lines). This prompts us to seek for a new way of plotting.
>>>>>>> added RMD

# ggplot

## What is ggplot?
In short, it's a plotting library that allows for more beautiful and complex plots than the default options in R. Below are some of the fundamental components of a ggplot:  
  
- ggplot() creates an empty coordinate graph. The idea is for you to add layers of graphs.
- Every function in ggplot2 takes a _mapping_ argument, which is always paired with _aes(x,y)_.
- **Aesthetic** = visual property of objects in the plot; _color, alpha, size, shape_
- **Facets** = subplots
- **Geom** = the geometrical object that a plot uses to represent data (e.g. bar, line)
  
_Code template_:
```{r eval=F}
ggplot(data = <DATA>) +  
  <GEOM_FUNCTION>(  
     mapping = aes(<MAPPINGS>),  
     stat = <STAT>,  
     position = <POSITION>  
  ) +  
  <COORDINATE_FUNCTION> +  
  <FACET_FUNCTION>  
```

## Aesthetics
```{r}
# Import libraries
library(ggplot2)
mpg # included in ggplot2
# Here, we map the aesthetic "color" to the variable "class"
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, color=class))
# Outside of aes(), the arguments are applied to the entire plot
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy), 
             color='blue', alpha=0.2, size=3, shape=18)
```

## Facets
_facet_wrap()_ lets you subplot by a single variable.  
To subplot by multiple variables, use _facet_grid()_.
```{r}
# face_wrap()
ggplot(data=mpg) +
  geom_point(mapping = aes(x=displ, y=hwy)) +
  facet_wrap(~ class, nrow=2)
# facet_grid()
ggplot(data=mpg) +
  geom_point(mapping = aes(x=displ, y=hwy)) +
  facet_grid(drv ~ cyl)
```

## Geoms
Some aesthetics work only with certain geoms (e.g. shape would not make sense with lines.)  
Global mapping is applied to each layer, and local maping is added on for that layer only.
```{r}
ggplot(data=mpg, mapping = aes(x=displ, y=hwy, color=drv)) +
  geom_smooth(mapping = aes(linetype=drv)) +
  geom_point()
```


# Geospatial Plotting
```{r}
```
