---
title: "Analytics 201: Session 1"
output: html_notebook
---

Welcome to WUDAC Analytics 201, Fall 2020! 

In this session, we will learn how to read in a dataset, clean it up, and manipulate it using a package called _dplyr_.


# Reading in data

We will be working with a wine dataset from the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/index.php). Let's see how to read it into your R workspace.

[Click here](https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv) to download the file containing the data.

```{r}
# First, move the the file to your desired folder location.
# Then, let's tell R where to find the file.

getwd() # this calls the current working directory

# If this is not where you want to save your work, change your path using the command below:
#setwd("C:/your/desired/path")
```
  
Once the file is in the right location, we can read in the dataset.
```{r}
# The file extension is .csv, so we use the classic read.csv() command
data <- read.csv('https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv', sep=';')

### Notice the "sep=';'" argument? That tells R to recognize ";" as the delimiter (typically, the delimiter is a comma, and we don't have to specify it.)
```

Now we have our data stored in a variable, so we are ready to clean it up.